Preliminary Thesis Defence

### 2014-01-06 south-africa.tex
conclusion

The other deviations are of a more practical nature.
Firstly, the data from the \textcite{saldru2008nids,saldru2012nids,saldru2013} surveys contains actual information on income,
including pension recipient status, whereas Duflo uses age as a proxy for recipient status.
Secondly, another minor deviation is the usage of \textcite[WHO Child Growth Standards]{who2006child},
in stead of \textcite[CDC Growth Charts: United States]{nchs2000cdc}, since these have superseded the CDC charts.
As long as all observations are held against the same standards, this should not be of any consequence.

@@

\subsubsection{Models and Variations}
We define the variables for our estimation equations.
The outcome variable is $y_{it}$, this outcome variable takes the form of the of the z-scores,
such as HAZ or WAZ.
Where $t$ denotes the time and $i$ the individual.
The individual and time fixes effects are denoted by $\gamma_i$ and $\lambda_t$ respectively.
Dummies for living in a household with a female or a male pension recipient are included as $P^f_{it}$ and $P^m_{it}$ respectively.
The dummy variable $T_{it}$ denoted the treatment status.
Lastly, $\epsilon_{it}$ is the error term, which is assumed to be distributed as:
\[
\epsilon_{it} \sim N(0,\sigma)
\]

@@

% CONCLUSIONS
% which hold? what do we think?

>>>>

The other deviations are of a more practical nature.
Firstly, the data from the \textcite{saldru2008nids,saldru2012nids,saldru2013nids} surveys contains actual information on income,
including pension recipient status, whereas Duflo uses age as a proxy for recipient status.
Secondly, another minor deviation is the usage of \textcite[WHO Child Growth Standards]{who2006child},
in stead of \textcite[CDC Growth Charts: United States]{nchs2000cdc}, since these have superseded the CDC charts.
As long as all observations are held against the same standards, this should not be of any consequence.

@@

\subsubsection{Models and Variations}
\label{sa:models}
We define the variables for our estimation equations.
The outcome variable is $y_{it}$, this outcome variable takes the form of the of the z-scores,
such as HAZ or WAZ.
Where $t$ denotes the time and $i$ the individual.
The individual and time fixes effects are denoted by $\gamma_i$ and $\lambda_t$ respectively.
Dummies for living in a household with a female or a male pension recipient are included as $P^f_{it}$ and $P^m_{it}$ respectively.
The dummy variable $T_{it}$ denoted the treatment status.
Lastly, $\epsilon_{it}$ is the error term, which is assumed to be distributed as:
\[
\epsilon_{it} \sim N(0,\sigma)
\]

@@

In our estimations we use the Difference-in-Differences method,
which is operationalised a fixed-effect panel estimation with a time effect.
The contruct three Right-Hand Side (RHS) model.
The first model is a fixed-effect estimation without the treatment dummy.
The second model is a fixed-effects estimation with the treatment dummy (\textbf{elig.men.60}).
the third model is a fixed-effects estimation with the treatment dummy,
as well as an interaction term with the male pension recipient dummy (\textbf{w\_spen\_m}).

On the Left-Hand Side (LHS) we use four different anthropometric z-scores,
Height-for-Age (HAZ), Weight-for-Age (WAZ), Weight-for-Height (WHZ), and BMI (BMIZ).
The first two z-scores are age based (ABZ), the latter two height based (since height factors in the BMI equation, HBZ).

Combining the three RHS models with the four LHS z-scores gives us twelve estimation equations.
In the estimation of these twelve equations, we find two effects.

Firstly, we find that in for both the Age-Based Z-scores (ABZs),
the treatment dummy is significant for the models 2 and the models 3
(we do not find this in model 1, since it is not included the model 1 equation).
In all cases, the coefficient for the treatment dummy is negative.
Since both these Age-Based Z-scores (ABZs) reflect long-term or past health issues,
this seems to indicate a negative correlation between the policy change on the long-term anthropometrics.
Since this is the treatment dummy, this would suggest a negative causal effect of male pension receipts on child health.
However, as discussed in \autoref{sa:results},
this is likely to be a consequence of the common trend not having been established.
Since we also observe a sharp drop in the number of children living with female pension recipients,
despite the fact that the female pension recipient base in unaffected by the policy change.

Secondly, we find that living with male pension recipient (\textbf{w\_spen\_m}),
is negatively correlated with a the BMI z-scores (\textbf{BMIZ}).
As discussed in \autoref{sa:results}, the Body-Mass-Index Z-score is a convex mapping of the Weight-for-Height z-score.
This can explain the fact that these essentialy similar estimations, give different significance levels.
Considering that the use of the Body Mass Index is far more wide spread than the Weight for Height metric,
we believe that the significance of these coefficients carries more weight.
The fact that these coefficients are negative, indicates that for this short-term z-score,
the prefered recipient would be female household members.

In conclusion, we find a negative correlation between the treatment dummy and the long-term ABZs.
This is likely to be a consequence of the common trend not having been established,
since there is also a drop in the number of children living with female pension recipients, which serves as the control group.
Furthermore, we find a negative correlation between the male pension recipient variable (\textbf{w\_spen\_m}) and the short-run \textbf{BMIZ}.

The later result seems to indicate that, at least in the short-run,
the prefered recipient of cash transfers would be a female household member.
As discussed, the selection bias issue in age,
leads to the need for a careful interpretation of the possible external validity of this results.

Regarding the former result, the significant negative coefficient of the treatment variable,
the analysis in \autoref{sa:results} seems to indicate a need to establish the common trend,
at least over the period 2008-2012.
Since the female recipient base in this period serves as a control group, we possibly could establish this, using the evolution here.
Alternatively, it would be possible to establish a trend using statistics on nationwide pension recipients.
